Busch, Khakhalin. Intrinsic temporal tuning

### 2018-12-08 reviews and responses.md

The number of degrees of freedom in question comes from a full linear model, and does mean that we treated different conditions as independent. Typically, to compare the numer of spikes in two conditions we used a 3-way ANCOVA (injection amplitude and duration as ordinal variables; group id as a nominal variable) with interactions, and repeated measures (wieh cell id included as a nominal variable). For example, for a comparison of "Control" group to "Sound" group we used the following R code (see `dynamic_spikes_plot.R` for the full code):

```R
ds = subset(d,is.element(Group,c("Control","Sound")))
summary(aov(data=ds,Spikes~Shape*Amp + Shape*Group + Amp*Group + Cell*Shape*Amp))
```

We now included a brief description of our model in the "Results" section; improved the description in the "Methods" section; and added several referrals to the "Methods" section in the "Results" section, to hint that the reader may find full information there. In particular, in the sentence that reports the first F-value for this study, we now added the following commentary:

"F(1,677)=30.4, p=5e-8, n cells=28, 29. Here and below, when F-values are reported, we use 4-way fixed effects analysis of variance with selected interactions, and cell id included as repeated measures factor, see "Methods" for a detailed description."

@@

> Page 6, line 22: "homologous to superior colliculus"

> Page 7, line 3: after multi sensory stimulation, "tectal neurons were more excitable than after visual stimulation alone). Many issues. The average number of spikes is reported..is this across all conductance curve conditions or just one of them? Next, in Fig 1E, "sync" and "flash" and "looming" all look pretty similar to each other. What are the degrees of freedom? Again, they seem very high like each condition might be being treated as independent.

> More stats: Are these multiple comparisons across conditions done in an unbiased manner, like with a Bonferroni correction or an ANOVA? It seems not.

1) As described above, we used a family of similar generalized linear models, followed by F-test for the analysis of variance, to compare across conditions. This type of analysis may be loosely referred to as ANOVA, although most textbooks would not recommend referring our models as an "ANOVA", as it included two ordinal variables, and a factor to account for repeated measures.

2) It is true that in our analysis we compared several pairs of conditions: Flash to Control, Looming to Control, Flash to Looming, Sound to Control, Flash to Sync, and Flash to Async. However, all of these comparisons were easily interpretable, as they form a logical progression of stimuli. These planned comparisons consistuted a small share of all possible post-hoc comparisons between groups that one could have run on this data (6 out of 30 possible pairs). Typically, for planned comparisons, researchers do not use an adjustment for multiple comparisons.

3) Finally, in this paper we report exact numbers all p-values, instead of indicating whether they were above or below any arbitrary threshold. This gives the reader the freedom to interpret our data against any significance threshold of their choice. As expected, some p-values were larger, and some were smaller, but we are confident that our overall conclusions are reasonable, expecially if comparing different pieces of evidence across the entire study.

Together, this makes us reasonably confident that a Bonferroni correction would be unnecessarily punitive for this type of analysis.

@@

> Page 8, lines 2-12. "Average firing rate" is a more common term than "spikiness". How about "gain" for amplitude tuning? I see that the authors are using the parameter a to indicate the non-linearity in the tuning curve. For "temporal tuning", why not use "adaptation index" or something like that, that could be plotted in Figure 1F? The "temporal tuning" is a measure of non-linearity, but it is not very intuitive. (It is not "incorrect" as the authors have it now, but I think it would be hard for the interdisciplinary readership of J Neuroscience to get on board easily.)

---------- TODO ----------- : Argue that average firing rate is usually applied to actual firing rate (either long-term, or instaneous) in realistic situations (in-vivo, or during spontaneous activity), while we look at a rather artificial excitability index. Gain and adaptation index should be included in the text, when "tunings" are introduced, but I prefer retaining original "tunings" for repeated use.

@@

> Figure 2: Units needed for Y axis on A, B (units of a are spikes per conductance squared and units of amplitude tuning are spikes/conductance)

---------- TODO ----------- : This is not true, both parameters are dimensionless, as both conductance and amplitude are treated as ordinal values. Reiterate this.

@@

> Page 8, about line 9: It is false that the numerical values are not interpretable, they are in an equation and have units.

We now replaced "not interpretable" with "not easily interpretable". We wanted our readers to know that a comparison of these values can be interpreted very easily, while exact numbers (say, 0.82) do not carry that much narrative weight.

@@

> Page 8, line 16-19: Cohen's D tells the discriminability rather than whether there is a difference between two quantities. A test like the Hoteling-2 test can provide significance measures for whether the means of the data in 2C are significant.

We report Cohen's d values as a measure of effect sizes, not as a quantification of significance.

---------- TODO ----------- : try Hoteling t-squared test

@@

> Figure 2B and Figure 3A, 3D don't show any post-hoc comparisons among the groups. Which groups differ? I see it in the text but it is hard to take it home.

---------- TODO ----------- : Add asterisks to figures?

> Figure 3E and G: how can we reconcile these accounts? In each of the groups (3G), there is at best a very weak relationship between temporal tuning and synaptic duration. Somehow, if you take the means of all of these groups, you can fit a line through them (Figure 3E). Why is that the right thing to do? Aren't the field of individual points (partially-transparent background) and the data plotted in 3G the relative quantities?

---------- TODO ----------- : Perhaps this will be more clear now, as the Simpson's praradox paragraph is back. Maybe we shoudl also add 1-2 more interpretative sentences to the Simpsons paradox paragraph? Also, do we need a reference there?

@@

> Figure 3C shows that monosynaptic amplitude and synaptic duration are correlated, as one would expect from the definition of synaptic duration (center of mass). Basically, as the monosynaptic amplitude gets bigger, the center of mass will shift earlier and earlier. So what independent information is really being conveyed in the synaptic duration? The correlation in 3C is so strong, I don't see that it is truly an independent quantity that teaches us something beyond the monosynaptic input strength. The monosynaptic input strength is much more defined, why not use that for examining correlation with temporal tuning (or adaptation index), average firing rate, and gain. The "synaptic duration" isn't really a measure of synaptic duration, which makes Page 10, lines 3-23 really hard to interpret.

On the correlation between monosynaptic amplitude and synaptic duration: this is a very valid observation, and it is also exactly what we claim in the text:

"As expected, cells with strong monosynaptic inputs had shorter synaptic currents, while polysynaptic activity made synaptic currents longer"

---------- TODO ----------- : Add a paragraph to the paper, explaining that synaptic duration is necessary, as it is what allows us to compare apples to apples, as we fit the synaptic currents exactly with a waveform we used to assess temporal tuning in dynamic clamp experiments. Then reiterate, and cite this paragraph here.

@@

> Figure 3G, page 10 lines 24-40. Creation of "super groups" is totally bogus. This part should be dropped. I see no strong significant relationship between temporal tuning and synaptic duration, and am not convinced synaptic duration is a meaningful quantity.

---------- TODO ----------- : See whether it is possible to use a different method here. The goal is to illustrate the fact that the weak correlation observed on full data is entirely driven by 3 "fast" datasets. Would some sort of a Tukey-like test work here?

@@

> I am very confused by Figure 4. Is it supposed to tell us that average firing rate is correlated with sodium channel conductances? Why is this experiment being done? What do we learn about the brain? I mean this genuinely, I feel I do not understand what the authors are trying to say that is novel. Of course we would expect such a relationship.

---------- TODO ----------- : Change the description; bring attention to the fact that the correlation exists (as predicted), but is also very bad (surprisinigly bad, in fact). Or maybe indeed drop it, and replace it with models predicting temporal and

@@

> Why not show the 8 parameter model fit of the current tuning curves?

---------- TODO ----------- : Do they mean a correlation-like plot that has model projection on X and estimations on Y? That would be interesting: do it.

@@

> Page 11, "the mechanisms behind temporal intrinsic plasticity". I'm confused. This section and Figure 4 seems to focus on average firing rate / spikiness rather than the temporal tuning parameter. So why the heading?

---------- TODO ----------- : The section is mostly about mechanisms, but Fig 4 for now is not; once we introduce new fit-figures, we'll be able to claim it.

### 2018-12-09 reviews and responses.md

> Page 8, about line 9: It is false that the numerical values are not interpretable, they are in an equation and have units.

We now replaced "not interpretable" with "not easily interpretable". We wanted our readers to know that a comparison of these values can be interpreted very easily, while exact numbers (say, 0.82) do not carry that much narrative weight.

> Page 8, line 16-19: Cohen's D tells the discriminability rather than whether there is a difference between two quantities. A test like the Hoteling-2 test can provide significance measures for whether the means of the data in 2C are significant.

We initially reported Cohen's d values as a measure of effect sizes, not as a quantification of significance. Now we also ran Hotelling t-squared tests on 2D sets of points, representing amplitude and temporal tuning values for different cells.

@@

"As expected, cells with strong monosynaptic inputs had shorter synaptic currents, while polysynaptic activity made synaptic currents longer"

---------- TODO ----------- : Add a paragraph to the paper, explaining that synaptic duration is necessary, as it is what allows us to compare apples to apples, as we fit the synaptic currents exactly with a waveform we used to assess temporal tuning in dynamic clamp experiments. Then reiterate, and cite this paragraph here.

> Figure 3G, page 10 lines 24-40. Creation of "super groups" is totally bogus. This part should be dropped. I see no strong significant relationship between temporal tuning and synaptic duration, and am not convinced synaptic duration is a meaningful quantity.

---------- TODO ----------- : Explain that essentially we are talking about strongly reshaped vs. weakly reshaped protocols. As we do not have apriory information about what protocols are inherently stronger, we can try to use Fig 2D to separate them into subgroups: weakly affected (Control, Sound, Looming) and storngly reshaped (Flash & both multisensory). Explain that it was the reason for creating "supergroups". Also downplay "supergroups", and remove this word. Make it more casual.

### 2018-12-09 reviews and responses.md

> Page 8, about line 9: It is false that the numerical values are not interpretable, they are in an equation and have units.

We now replaced "not interpretable" with "not easily interpretable". We wanted our readers to know that a comparison of these values can be interpreted very easily, while exact numbers (say, 0.82) do not carry that much narrative weight.

> Page 8, line 16-19: Cohen's D tells the discriminability rather than whether there is a difference between two quantities. A test like the Hoteling-2 test can provide significance measures for whether the means of the data in 2C are significant.

We initially reported Cohen's d values as a measure of effect sizes, not as a quantification of significance. Now we also ran Hotelling t-squared tests on 2D sets of points, representing amplitude and temporal tuning values for different cells, across different groups.

@@

---------- TODO ----------- : Add a paragraph to the paper, explaining that synaptic duration is necessary, as it is what allows us to compare apples to apples, as we fit the synaptic currents exactly with a waveform we used to assess temporal tuning in dynamic clamp experiments. Then reiterate, and cite this paragraph here.

> Figure 3G, page 10 lines 24-40. Creation of "super groups" is totally bogus. This part should be dropped. I see no strong significant relationship between temporal tuning and synaptic duration, and am not convinced synaptic duration is a meaningful quantity.

---------- TODO ----------- : Explain that essentially we are talking about strongly reshaped vs. weakly reshaped protocols. As we do not have apriory information about what protocols are inherently stronger, we can try to use Fig 2D to separate them into subgroups: weakly affected (Control, Sound, Looming) and storngly reshaped (Flash & both multisensory). Explain that it was the reason for creating "supergroups". Also downplay "supergroups", and remove this word. Make it more casual.

Or maybe indeed just drop it. Leave two sentences: one listing r values, and another one mentioning that the negative correlation that exists for full data is entirely driven by these 3 groups, but not by those other 3 groups.

### 2018-12-09 main.tex
responses: minor notes on stats

To visualize and interpret differences in neuronal tuning, we quantified each of the three aspects of intrinsic tuning (average spikiness, amplitude tuning, and temporal tuning) with one value per neuron (see Methods). We used the mean number of spikes across all conditions as the measure of "spikiness"; the linear slope of the number of spikes as a function of conductance amplitude as the measure of "amplitude tuning", and the quadratic term of the curvilinear regression for the number of spikes as a function of input duration as the value of "temporal tuning" (Figure 2A). The numerical values of these "tuning coefficients" are not interpretable, but they capture the character of tuning curves for each neuron (Figure 2A). All three parameters differed across experimental groups: F(5,160)=3.1, p=0.01 for average spikiness (Figure 2C; see "Methods" for model description); F(5,160)=4.8, p=4e-4 for amplitude tuning (Figure 2B); and F(5,160)=3.6, p=4e-3 for temporal tuning (Figure 2C).

A visual comparison of neuronal tuning in different experimental groups (Fig 2C) shows that acoustic stimulation had opposite effects when provided on its own (without visual stimulation) than when added to visual flashes. Compared to control neurons, cells exposed to sound had lower amplitude tuning coefficients (flatter curve, Cohen’s d=$-$0.29), and more negative (curving down, d=$-$0.27) temporal tuning coefficients. In contrast, when sounds were added to flashes, both transfer functions became less flat (d=0.65 and 0.26 for "Async" compared to “Flash”, for amplitude and temporal tuning respectively), and more like curves for control neurons. Thus, acoustic stimulation tended to tune the network in the same direction as visual stimulation when delivered alone (Figure 2D), but negated the effect of visual stimulation when combined with it. This may imply that multisensory integration in the tectum is dominated by inhibition (see Discussion). Note also that looming stimuli had a weaker effect on neuronal tuning compared to flashes, both in terms of amplitude (d=0.44) and temporal tuning (d=0.22).

@@

%%% --- This section was commented out in all official submissions

@@

To better understand the interplay between the stimulation-induced retuning of the entire tectal network, and tuning of individual cells within each network, we looked at the interaction of intrinsic and synaptic temporal properties separately for each treatment group (Figure 3G). The co-tuning seemed present in Control, Looming, and Sound groups (r=$-$0.11, $-$0.36, and $-$0.47), but was absent in all groups that experienced flashes in any form (r=0.27, $-$0.01, and 0.01 for Flash, Sync, and Async groups respectively). If three slow, and three fast (Flash-like) groups were combined into two larger "supergroups", the Control-Looming-Sound neurons exhibited a homeostatic negative correlation between synaptic duration and intrinsic temporal tuning (r=$-$0.29, p=0.01, n=70), while the Flash-Sync-Async neurons lacked cells preferring slow stimuli, and so lacked the co-tuning (r=0.08, p=0.50, n=80). This suggests that when fast visual stimulation made the tectal network selective to short activation patterns, it happened not because all neurons uniformly changed their properties, but because a particular subset of neurons, namely those that receive strong direct inputs from the retina and weak recurrent inputs, become less excitable. Normally these neurons experience fast synaptic inputs, but are most responsive to slow conductances; in overstimulated networks however they became selective to fast conductances.

@@

Knowing that tectal neurons can tune to inputs of different temporal dynamics, we then investigated the cellular mechanisms that could underlie this tuning. For each cell, we used a sequence of voltage steps (Figure 4A) to activate Na and K conductances, and quantified ionic current amplitudes and activation potentials (Figure 4B) as it was done in earlier studies \citep{ciarleglio2015}. Together with cell membrane resistance (Rm) and capacitance (Cm) it gave us eight intrinsic parameters for every cell: peak amplitudes for sodium current, early (transient) potassium current, and late (stable) potassium current (INa, IKt, IKs respectively), and activation potentials for these three currents (VNa, VKt, and VKs).

>>>>

To visualize and interpret differences in neuronal tuning, we quantified each of the three aspects of intrinsic tuning (average spikiness, amplitude tuning, and temporal tuning) with one value per neuron (see Methods). We used the mean number of spikes across all conditions as the measure of "spikiness"; the linear slope of the number of spikes as a function of conductance amplitude as the measure of "amplitude tuning", and the quadratic term of the curvilinear regression for the number of spikes as a function of input duration as the value of "temporal tuning" (Figure 2A). The numerical values of these "tuning coefficients" are not easily interpretable, but they capture the character of tuning curves for each neuron (Figure 2A). All three parameters differed across experimental groups: F(5,160)=3.1, p=0.01 for average spikiness (Figure 2C; see "Methods" for model description); F(5,160)=4.8, p=4e-4 for amplitude tuning (Figure 2B); and F(5,160)=3.6, p=4e-3 for temporal tuning (Figure 2C).

A visual comparison of neuronal tuning in different experimental groups (Fig 2C) shows that acoustic stimulation had opposite effects when provided on its own (without visual stimulation) than when added to visual flashes. Compared to control neurons, cells exposed to sound had slightly lower amplitude tuning coefficients (flatter curve, Cohen’s d=$-$0.29), and more negative (curving down, d=$-$0.27) temporal tuning coefficients (Hotelling t-squared test p=0.2). When sounds were added to flashes, however, both transfer functions became less flat (d=0.65 and 0.26 for "Async" compared to “Flash”, for amplitude and temporal tuning respectively), and more like curves for control neurons (Hotteling test p=0.04). Thus, acoustic stimulation tended to tune the network in the same direction as visual stimulation when delivered alone (Figure 2D), but negated the effect of visual stimulation when combined with it. This may imply that multisensory integration in the tectum is dominated by inhibition (see Discussion). Note also that looming stimuli seemed to have weaker effects on neuronal tuning compared to flashes, both in terms of amplitude (d=0.44) and temporal tuning (d=0.22).

@@

%%% --- This section below was commented out in prior submissions

@@

To better understand the interplay between the stimulation-induced retuning of the entire tectal network, and tuning of individual cells within each network, we looked at the interaction of intrinsic and synaptic temporal properties separately for each treatment group (Figure 3G). The co-tuning seemed present in Control, Looming, and Sound groups (r=$-$0.11, $-$0.36, and $-$0.47), but was absent in groups that experienced flashes, and so were subjected to stronger retuning (r=0.27, $-$0.01, and 0.01 for Flash, Sync, and Async groups respectively). If three slow, and three fast (Flash-like) groups were combined into two larger "supergroups", the Control-Looming-Sound neurons exhibited a homeostatic negative correlation between synaptic duration and intrinsic temporal tuning (r=$-$0.29, p=0.01, n=70), while the Flash-Sync-Async neurons lacked cells preferring slow stimuli, and so lacked the co-tuning (r=0.08, p=0.50, n=80). This suggests that when fast visual stimulation made the tectal network selective to short activation patterns, it happened not because all neurons uniformly changed their properties, but because a particular subset of neurons, namely those that receive strong direct inputs from the retina and weak recurrent inputs, become less excitable. Normally these neurons experience fast synaptic inputs, but are most responsive to slow conductances; in overstimulated networks however they became selective to fast conductances.

@@

Knowing that tectal neurons can tune to inputs of different temporal dynamics, we then tried to identify the cellular mechanisms underlying this tuning. For each cell, we used a sequence of voltage steps (Figure 4A) to activate Na and K conductances, and quantified ionic current amplitudes and activation potentials (Figure 4B) as it was done in earlier studies \citep{ciarleglio2015}. Together with cell membrane resistance (Rm) and capacitance (Cm) it gave us eight intrinsic parameters for every cell: peak amplitudes for sodium current, early (transient) potassium current, and late (stable) potassium current (INa, IKt, IKs respectively), and activation potentials for these three currents (VNa, VKt, and VKs).
